---
output:
  html_document: default
  pdf_document: default
---
**Nama: Riofebri Prasetia** <br>
**NIM: 221911192** <br>
**Kelas: 3SI1** <br>
**Tugas pertemuan 9** <br>

## 9.10

```{r}
#The correlation matrix for chicken-bone measurements:
chickenBon <- matrix(c(1, 0.505, 0.569, 0.602, 0.621, 0.603,
                 0.505, 1, 0.422, 0.467, 0.482, 0.45,
                 0.569, 0.422, 1, 0.926, 0.877, 0.878,
                 0.602, 0.467, 0.926, 1, 0.874, 0.894,
                 0.621, 0.482, 0.877, 0.874, 1, 0.937,
                 0.603, 0.45, 0.878, 0.894, 0.937, 1),ncol = 6, nrow = 6)
chickenBon
```
```{r}
#The following estimated factor loadings were extracted by the maximum likelihood procedure:
variabel <- c("1. Skull length",
                       "2. Skull breadth",
                       "3. Femur length",
                       "4. Tibia length",
                       "5. Humerus length",
                       "6. Ulna length")
F1 <- c(0.602,
        0.467,
        0.926,
        1,
        0.874,
        0.894)
F2 <- c(0.2, 0.154, 0.143, 0, 0.476, 0.327)
F_1 <- c(0.484, 0.375, 0.603, 0.519, 0.861, 0.744)
F_2 <- c(0.411, 0.319, 0.717, 0.855, 0.499, 0.594)
table <- data.frame(variabel, F1, F2, F_1, F_2)
table
```
Using the unrotated estimated factor loadings, obtain the maximum likelihood estimates of the following.
(a) The specific variances.
```{r}
# Specific Variance F estimated

matrixSpi <- matrix(c(0), nrow = nrow(chickenBon), ncol = ncol(chickenBon))
SpiChickenBon <- 1 - (table$F1^2 + table$F2^2)
for(i in 1:ncol(chickenBon))
{
        matrixSpi[i,i] <- SpiChickenBon[i]
}
matrixSpi

# Specific variance F* varimax rotated estimated
matrixSpiF <- matrix(c(0), nrow = nrow(chickenBon), ncol = ncol(chickenBon))
SpiChickenBonF <- 1 - (table$F_1^2 + table$F_2^2)
for(i in 1:ncol(chickenBon))
{
        matrixSpiF[i,i] <- SpiChickenBonF[i]
}
matrixSpiF
```

(b) The communalities. 
```{r}
# The Communalities F estimated
hChickenBon <- table$F1^2 + table$F2^2
hChickenBon

# The Communalities F* varimax rotated estimated
hChickenBonF <- table$F_1^2 + table$F_2^2
hChickenBonF

```

(c) The proportion of variance explained by each factor.
```{r}
# The proportion of variance explained by each factor F estimated
LChickenBon <- data.frame(c(table[2], table[3]))
LChickenBon
proChickenBon <- (table$F1^2 + table$F2^2)/nrow(LChickenBon)
proChickenBon

# The proportion of variance explained by each factor F* varimax rotated estimated
LChickenBonF <- data.frame(c(table[4], table[5]))
LChickenBonF
proChickenBonF <- (table$F_1^2 + table$F_2^2)/nrow(LChickenBon)
proChickenBonF

```

(d) The residual matrix $R - \hat{L_{z}}\hat{L^{'}_{z}} - \hat{\Psi_{z}}$
```{r}
LChickenBon <- as.matrix(LChickenBon)
LChickenBonF <- as.matrix(LChickenBonF)

# The Residual Matrix F estimated
matrixChickenBon <- chickenBon - LChickenBon%*%t(LChickenBon) - matrixSpi
matrixChickenBon

# The Residual Matrix F* varimax rotated estimated
matrixChickenBonF <- chickenBon - LChickenBonF%*%t(LChickenBonF) - matrixSpiF
matrixChickenBonF

```



## 9.19

A firm is attempting to evaluate the quality of its sales staff and is trying to find an examination or series of tests that may reveal the potential for good performance in sales. The firm has selected a random sample of 50 sales people and has evaluated each on 3 measures of performance: growth of sales, profitability of sales, and new-account sales.These measures have been converted to a scale, on which 100 indicates "average" performance. Each of the 50 individuals took each of 4 tests, which purported to measure creativity, mechanical reasoning, abstract reasoning, and mathematical ability, respectively. The n = 50 observations on p = 7 variables are listed in Table 9.12.

*Note*: The components of x must be standardized using the sample means and variances calculated from the original data.

```{r}
url <- "https://raw.githubusercontent.com/rii92/tugas-APG/main/tugas%20pertemuan%209/dataSoal9.19.csv"
salesPeople <- read.csv(url)
salesPeople <- subset(salesPeople, select = -c(Salesperson))
salesPeople
```
**x1: Sales growth**<br> 
**x2: Sales Profit ability**<br> 
**x3: New account sales**<br> 
**x4: Creativity test**<br> 
**x5: Mechanical reasoning test**<br> 
**x6: Abstract reasoning test**<br> 
**x7: Mathematics test**<br> 

(a) Assume an orthogonal factor model for the standardized variables $Z_{i} = (X_{i} - \mu_{i})/ \sigma_{ii}^{1/2}$, i = 1, 2, ... , 7. Obtain either the principal component solution or the maximum likelihood solution for m = 2 and m = 3 common factors. 
```{r}
factor2 <- factanal(x = salesPeople, factors = 2)
factor3 <- factanal(x = salesPeople, factors = 3)
factor2
factor3
```
(b) Given your solution in (a), obtain the rotated loadings form = 2 and m = 3. Compare the two sets of rotated loadings. Interpret the m = 2 and m = 3 factor solutions. 
Jawab:
metode yang digunakan adalah maximum likelihood solution, pada faktor m = 2 untuk faktor pertama memiliki faktor untuk semua variabel diatas 0.5 kecuali untuk variabel creativity test, untuk faktor kedua hampir setengah dari variabel yang memiliki nilai faktor estimasi sebesar di bawah 0.5. Pada faktor m = 3 untuk faktor pertama terdapat variabel Creativity test dan abstract reasoning test dengan faktor di bawah 0.5, faktor kedua terdapat variabel New account sales dan Creativity test dengan faktor estimasi di atas 0.5, untuk faktor ketiga terdapat variabel Abstract reasoning test dengan faktor estimasi di atas 0.5. dalam hal ini berdasarkan faktor estimasi maka untuk faktor m = 2 sudah bisa mewakili karena sebagian besar variabel faktor yang memiliki nilai di atas 0.5 atau bernilai besar dengan proporsi yang seimbang

(c) List the estimated communalities, specific variances, and $\hat{L}\hat{L^{'}} + \hat{\Psi}$ for the m = 2 and m = 3 solutions. Compare the results. Which choice of m do you prefer at this point? Why? 
```{r}
#untuk m = 2
communalities2 <- matrix(c(0), nrow = nrow(factor2$loadings), ncol = 1)
SpecificVariance2 <- matrix(c(0), nrow = nrow(factor2$loadings), ncol = nrow(factor2$loadings))
for(i in 1:nrow(factor2$loadings))
{
        communalities2[i] <- (factor2$loadings[i,1])^2 + (factor2$loadings[i,2])^2
}

for(i in 1:nrow(factor2$loadings))
{
        SpecificVariance2[i,i] <- 1 - communalities2[i]
}


communalities2
SpecificVariance2
factor2$loadings%*%t(factor2$loadings) + SpecificVariance2
cor(salesPeople)
```
```{r}
#untuk m = 3
communalities3 <- matrix(c(0), nrow = nrow(factor3$loadings), ncol = 1)
SpecificVariance3 <- matrix(c(0), nrow = nrow(factor3$loadings), ncol = nrow(factor3$loadings))
for(i in 1:nrow(factor3$loadings))
{
        communalities3[i] <- (factor3$loadings[i,1])^3 + (factor3$loadings[i,3])^3
}

for(i in 1:nrow(factor3$loadings))
{
        SpecificVariance3[i,i] <- 1 - communalities3[i]
}


communalities3
SpecificVariance3
factor3$loadings%*%t(factor3$loadings) + SpecificVariance3
cor(salesPeople)
```
Interpretasi:
pada communalities terlihat bahwa untuk faktor m = 2 menunjukkan bahwa kedua faktor tersebut merupakan persentase yang besar dari varians sampel setiap variabel dibanding faktor m = 3. Kemudian pada hasil $LL^{'} + \Psi$ pada faktor m = 2 hampir mendekati dengan matriks korelasi R dibanding dengan faktor m = 3. Dalam hal ini, faktor m = 2 masih unggul di banding faktor m = 3.

(d) Conduct a test of H0: $LL^{'} + \Psi$ + '\}I versus H1: $\Sigma \ne LL^{'} + \Psi$ for both m = 2 and m = 3 at the a = .01 level. With these results and those in Parts b and c, which choice of $m$ appears to be the best? 
```{r}
#test dengan alpha = .01
factor2$PVAL
factor3$PVAL
```
Interpretasi:
kedua kondisi (m = 2 dan m = 3) memiliki p-value kurang dari 0.01. Artinya kedua kondisi masih memiliki hasil $LL^{'} + \Psi$ yang sama dengan covariance matrix $\Sigma$. Dibanding kedua kondisi maka akan diambil p-value yang paling kecil untuk kondisi faktor yang terbaik yaitu faktor m = 2

(e) Suppose a new salesperson, selected at random, obtains the test scores 
$x^{'} = [x_{1}, x_{2},... , x_{7}]$ = [ 110, 98, 105, 15, 18, 12, 35] . Calculate the salesperson's factor score using the weighted least squares method and the regression method.
```{r}
z_test <- c(110, 98, 105, 15, 18, 12, 35)

#untuk m=2
# Weighted least squares
L2 <- factor2$loadings%*%factor2$rotmat
f2 <- solve(t(L2)%*%solve(SpecificVariance2)%*%L2)%*%t(L2)%*%solve(SpecificVariance2)%*%z_test
print(paste("Weighted least squares (m=2):"))
f2

#Regression
f2reg <- t(L2)%*%solve(cor(salesPeople))%*%z_test
print(paste("Regression (m=2):"))
f2reg

#untuk m=3
# Weighted least squares
L3 <- factor3$loadings%*%factor3$rotmat
f3 <- solve(t(L3)%*%solve(SpecificVariance3)%*%L3)%*%t(L3)%*%solve(SpecificVariance3)%*%z_test
print(paste("Weighted least squares (m=3):"))
f3

#Regression
f3reg <- t(L3)%*%solve(cor(salesPeople))%*%z_test
print(paste("Regression (m=3):"))
f3reg
```

## 9.20
Using the air-pollution variables X1, X2, X5, and X6 given in Table 1.5, generate the sample covariance matrix. 

```{r}
url <- "https://raw.githubusercontent.com/rii92/tugas-APG/main/tugas%20pertemuan%208/table%201.5.csv"
airPollution <- read.csv(url)
airPollution <- subset(airPollution, select = -c(co, NO, HC))
airPollution
```
(a) Obtain the principal component solution to a factor model with m = 1 and m = 2. 
```{r}
eigenAirPollution <- eigen(cov(airPollution))

F1AirPollution <- sqrt(eigenAirPollution$values[1])*eigenAirPollution$vectors[,1]
F2AirPollution <- sqrt(eigenAirPollution$values[2])*eigenAirPollution$vectors[,2]
print(paste("factor:"))
F1AirPollution
#untuk m = 1
LAirPollution1 <- F1AirPollution%*%t(F1AirPollution)
print(paste("L*t(L)"))
LAirPollution1
SpecificVariance1AirPollutionPC <- cov(airPollution) - LAirPollution1
print(paste("Specific variance"))
SpecificVariance1AirPollutionPC
```
```{r}
#untuk m = 2
df2 <- data.frame(F1AirPollution, F2AirPollution)
FAirPollution <- data.matrix(df2)
print(paste("Factor:"))
FAirPollution
LAirPollution2 <- FAirPollution%*%t(FAirPollution)
print(paste("L*t(L)"))
LAirPollution2
SpecificVariance2AirPollutionPC <- cov(airPollution) - LAirPollution2
print(paste("Specific Variance"))
SpecificVariance2AirPollutionPC
```

(b) Find the maximum likelihood estimates of **L** and $\Psi$ for m = 1 and m = 2. 
```{r}
factor1AirPollution <- factanal(x = airPollution, factors = 1)
#factor2AirPollution <- factanal(x = airPollution, factors = 2)

# m = 1
print(paste("untuk m = 1"))
factor1AirPollution$loadings
SpecificVariance1AirPollution <- matrix(c(0), ncol = nrow(factor1AirPollution$loadings), nrow = nrow(factor1AirPollution$loadings))

for(i in 1:nrow(factor1AirPollution$loadings))
{
        SpecificVariance1AirPollution[i,i] <- 1 - (factor1AirPollution$loadings[i,1])^2
}

SpecificVariance1AirPollution
```
```{r}
# m = 2
#print(paste("untuk m = 2"))
#factor2AirPollution$loadings
#SpecificVariance2AirPollution <- matrix(c(0), ncol = nrow(factor2AirPollution$loadings), nrow = nrow(factor2AirPollution$loadings))

#for(i in 2:nrow(factor2AirPollution$loadings))
#{
#        SpecificVariance2AirPollution[i,i] <- 1 - ((factor2AirPollution$loadings[i,2])^2 + (factor2AirPollution$loadings[i,1])^2)
#}

#SpecificVariance2AirPollution
```
(c) Compare the factorization obtained by the principal component and maximum likelihood methods.
<br>Interpretasi:
pada faktor untuk metode principal component solution pada m = 1 dan m = 2 terdapat pembeda antara faktor positif dan negatif (masih belum diketahui pembedanya apa), untuk specific variance secara garis besar tidak mendekati nilai specific variance karena selain matriks diagonal, nilainya tidak semua 0. Pada faktor untuk metode maximum likelihood pada m = 1 (untuk m = 2 tidak bisa dilakukan karena terlalu banyak untuk variabel sama dengan 4) juga terdapat pembeda faktor positif dan negatif, untuk specific variance hampir mendekati karena nilai selain matrix diagonal adalah 0. Dalam hal ini pada kasus ini sebaiknya menggunakan metode maximum likelihood untuk analisis faktor