---
title: "APG_Pertemuan_5"
author: "Riofebri Prasetia (221911192)"
date: "3/19/2022"
output: html_document
---

```{r library, message=FALSE}
library('MVN')
library('mvoutlier')
library('rcompanion')
library('caret')
library('RVAideMemoire')
library('ellipse')
library('qqplotr')
library('evd')
library('outliers')
library('ggplot2')
library('SciViews')
```

Soal
#6.1.
Construct and sketch a joint 95% confidence region for the mean difference 
vector B using the effluent data and results in Example 6.1. Note that the point 
B == 0 falls outside the 95% contour. Is this result consistent with the test of 
H0: B == 0 considered in Example 6.1? Explain.
BOD_Commercial_lab(xi) 
SS_Commercial_lab(x2) 
BOD_Statelabofhygiene(x3) 
SS_Statelabofhygiene(x4) 

```{r}
url1<-"https://raw.githubusercontent.com/rii92/tugas-APG/main/20%20Maret%202022/APG_dataNo_6.1.csv"
dataeffluent1 <- read.csv(url1)
dataeffluent2 <- data.frame(dataeffluent1$x1j1, dataeffluent1$x1j2, dataeffluent1$x2j1, dataeffluent1$x2j2)
colnames(dataeffluent2)[colnames(dataeffluent2)=="dataeffluent1.x1j1"] <-"x1j1"
colnames(dataeffluent2)[colnames(dataeffluent2)=="dataeffluent1.x1j2"] <-"x1j2"
colnames(dataeffluent2)[colnames(dataeffluent2)=="dataeffluent1.x2j1"] <-"x2j1"
colnames(dataeffluent2)[colnames(dataeffluent2)=="dataeffluent1.x2j2"] <-"x2j2"
dataeffluent2
```

```{r}
dj1 <- dataeffluent1$x1j1-dataeffluent1$x2j1
dj2 <- dataeffluent1$x1j2-dataeffluent1$x2j2
data_dj <- data.frame(dj1, dj2)
data_dj
```


##Membuat Ellipse
```{r}
mu <- c(0,0)
X <- data_dj
n <- nrow(X)
p <- ncol(X)
xbar <- colMeans(X)
S <- cov(X)
library(car)
alpha<-0.05
tconst <- sqrt((p/n)*((n-1)/(n-p)) *
                 qf(1-alpha,p,n-p))
id <- c(1,2)
plot(ellipse(center=xbar[id], shape=S[id,id],
             radius=tconst, draw=F),type="n", xlab = "Commercial Lab", ylab = "State lab of hygiene")
lines(ellipse(center=xbar[id], shape=S[id,id],
              radius=tconst, lwd=3))
points(mu[1],mu[2],col="red", pch=16)
text(mu[1]+0.75,mu[2],expression(d[0]))
text(xbar[1]+0.75,xbar[2],expression(bar(d)))
```
Interpretasi:
bisa dilihat bahwa B = (0,0) diluar daerah kepercayaan, dari ini bisa disimpulkan bahwa H0 : d' = 0 ditolak pada daerah kepercayaan 95%. Sehingga bisa kita simpulkan bahwa hasil data Effluent dengan uji H0 : B = 0 adalah konsisten.


#6.2.
Using the information in Example 6.1. construct the 95% Bonferroni simulta neous intervals for the components of the mean difference vector 8. Compare the lengths of these intervals with those of the simultaneous intervals constructed in the example.

##Interval Bonferroni
```{r}
bon <- NULL
alpha2 <- 1 - 0.05/(2*p)
for(k in 1:p){
 
 bon <- c(bon,
 xbar[k] - sqrt(S[k,k]/n) * qt(alpha2, df = n-1),
 xbar[k] + sqrt(S[k,k]/n) * qt(alpha2, df = n-1))
}
round(bon, p)
```

##Interval Simultan T-square test
```{r}
library(MVTests)
mu <- c(0, 0)
ujiT <- OneSampleHT2(data = X, mu0 = mu, alpha = 0.05)
summary(ujiT)
```
Interpretasi:
Hasilnya ialah interval T-square test lebih besar di banding interval Bonferroni

#6.3
The data corresponding to sample 8 in Table 6.1 seem unusually large. Remove 
sample 8. Construct a joint 95% confidence region for the mean difference vector B and the 95% Bonferroni simultaneous intervals for the components of the 
mean difference vector. Are the results consistent with a test of H0 : B = 0? 
Discuss. Does the "outlier" make a difference in the analysis of these data?

##Hapus data ke-8
```{r}
datadj_after_drop <- data_dj[-8,]
datadj_after_drop
```
##Membuat Ellipse
```{r}
mu <- c(0,0)
X <- datadj_after_drop
n <- nrow(X)
p <- ncol(X)
xbar <- colMeans(X)
S <- cov(X)
library(car)
alpha<-0.05
tconst <- sqrt((p/n)*((n-1)/(n-p)) *
                 qf(1-alpha,p,n-p))
id <- c(1,2)
plot(ellipse(center=xbar[id], shape=S[id,id],
             radius=tconst, draw=F),type="n", xlab = "Commercial Lab", ylab = "State lab of hygiene", xlim = c(-25,1))
lines(ellipse(center=xbar[id], shape=S[id,id],
              radius=tconst, lwd=3))
points(mu[1],mu[2],col="red", pch=16)
text(mu[1]+0.75,mu[2],expression(d[0]))
text(xbar[1]+0.75,xbar[2],expression(bar(d)))
```
##Interval Bonferroni
```{r}
bon <- NULL
alpha2 <- 1 - 0.05/(2*p)
for(k in 1:p){
 
 bon <- c(bon,
 xbar[k] - sqrt(S[k,k]/n) * qt(alpha2, df = n-1),
 xbar[k] + sqrt(S[k,k]/n) * qt(alpha2, df = n-1))
}
round(bon, p)
```

##Interval Simultan T-square test
```{r}
library(MVTests)
mu <- c(0, 0)
ujiT <- OneSampleHT2(data = X, mu0 = mu, alpha = 0.05)
summary(ujiT)
```

Interpretasi:
Berdasarkan hasil, pencilan tidak mempengaruhi hasil. Artinya data Effluent konsisten dengan uji H0:B = 0.

#6.4
Refer to Example 6.1. 
(a) Redo the analysis in Example 6.1 after transforming the pairs of observations to ln(BOD) and ln(SS). 
(b) Construct the 95% Bonferroni simultaneous intervals for the components of the mean vector B of transformed variables. 
(c) Discuss any possible violation of the assumption of a bivariate normal distribution for the difference vectors of transformed observations.

Melakukan transformasi
```{r}
x1j1_at <- log(dataeffluent1$x1j1)
x1j2_at <- log(dataeffluent1$x1j2)
x2j1_at <- log(dataeffluent1$x2j1)
x2j2_at <- log(dataeffluent1$x2j2)
data_effluent2_at <- data.frame(x1j1_at, x1j2_at, x2j1_at, x2j2_at)
data_effluent2_at
```
```{r}
dj1_at <- data_effluent2_at$x1j1_at-data_effluent2_at$x1j2_at
dj2_at <- data_effluent2_at$x2j1_at-data_effluent2_at$x2j2_at
data_dj_at <- data.frame(dj1_at, dj2_at)
data_dj_at
```
##Membuat Ellipse
```{r}
mu <- c(0,0)
X <- data_dj_at
n <- nrow(X)
p <- ncol(X)
xbar <- colMeans(X)
S <- cov(X)
library(car)
alpha<-0.05
tconst <- sqrt((p/n)*((n-1)/(n-p)) *
                 qf(1-alpha,p,n-p))
id <- c(1,2)
plot(ellipse(center=xbar[id], shape=S[id,id],
             radius=tconst, draw=F),type="n", xlab = "t(Commercial Lab)", ylab = "t(State lab of hygiene)", xlim = c(-1.4,0.25))
lines(ellipse(center=xbar[id], shape=S[id,id],
              radius=tconst, lwd=3))
points(mu[1],mu[2],col="red", pch=16)
text(mu[1]+0.05,mu[2],expression(d[0]))
text(xbar[1]+0.05,xbar[2],expression(bar(d)))
```
Interpretasi:
Hasilnya menunjukkan bahwa setelah dilakukan transformasi ln atau log(BOD) dan ln(SS), titik B = 0 berada di luar penerimaan H0. Maka data Effluent setelah transformasi ln konsisten dengan uji H0: B = 0. 

##Interval Bonferroni
```{r}
bon <- NULL
alpha2 <- 1 - 0.05/(2*p)
for(k in 1:p){
 
 bon <- c(bon,
 xbar[k] - sqrt(S[k,k]/n) * qt(alpha2, df = n-1),
 xbar[k] + sqrt(S[k,k]/n) * qt(alpha2, df = n-1))
}
round(bon, p)
```

##Interval Simultan T-square test
```{r}
library(MVTests)
mu <- c(0, 0)
ujiT <- OneSampleHT2(data = X, mu0 = mu, alpha = 0.05)
summary(ujiT)
```
Interpretasi:
Interval kepercayaan T-square masih lebih luas di banding interval bonferroni.


```{r}
pairs(data_dj_at)
```

```{r}
qqnorm(data_dj_at$dj1_at)
qqline(data_dj_at$dj1_at)
```

```{r}
qqnorm(data_dj_at$dj2_at)
qqline(data_dj_at$dj2_at)
```

```{r}
mvn(data_dj_at, multivariatePlot = "qq", showOutliers = TRUE)[4]
```

```{r}
mvn(data = data_dj_at[, c(1,2)], multivariatePlot = 'plot', showOutlier 
= TRUE)
```
```{r}
mshapiro.test(data_dj_at[, c(1,2)])
```

Interpretasi:
Pada test Multivariate normal Henze-Zirkler dan shapiro wilk menyatakan bahwa data multivariate/bivariate normal. Akan tetapi ketika melihat plot bivariate, sebaran data nya tidak terlihat pola arahnya. Demikian itu bisa disimpulkan belum cukup bukti untuk menyatakan bahwa data effluent setelah transformasi melanggar asumsi distribusi normal

#6.17
The data in Table 6.8 were collected to test two psychological models of numerical cognition. Does the processing of numbers depend on the way the numbers are presented (words, Arabic digits)? Thirty-two subjects were required to make a series of quick numerical judgments about two numbers presented as either two number words ("two," "four") or two single Arabic digits ("2," "4"). The subjects were asked to respond "same" if the two numbers had the same numerical parity (both even or both odd) and "different" if the two numbers had a different parity (one even, one odd). Half of the subjects were assigned a block of Arabic digit trials, followed by a block of number word trials, and half of the subjects received the blocks of trials in the reverse order. Within each block, the order of "same" and "different" parity trials was randomized for each subject. For each of the four combinations of parity and format, the median reaction times for correct responses were recorded for each 
subject. Here
X1 = median reaction time for word format-different parity combination 
X2 = median reaction time for word format-same parity combination 
X3 = median reaction time for Arabic format-different parity combination X4 = median reaction time for Arabic format-same parity combination

(a) Test for treatment effects using a repeated measures design. Set a = .05. 
(b) Construct 95% (simultaneous) confidence intervals for the contrasts representing the number format effect, the parity type effect and the interaction 
effect. Interpret the resulting intervals. 
(c) The absence of interaction supports the M model of numerical cognition, 
while the presence of interaction supports the C and C model of numerical 
cognition. Which model is supported in this experiment? 
(d) For each subject, construct three difference scores corresponding to the 
number format contrast, the parity type contrast, and the interaction

contrast. Is a multivariate normal distribution a reasonable population 
model for these data? Explain

```{r}
url <- "https://raw.githubusercontent.com/rii92/tugas-APG/main/20%20Maret%202022/APG_dataNo_6.17.csv"
datasoal_5 <- read.csv(url)
datasoal_5
```
```{r}
Cmat <- matrix(c(
 -1, 1, -1, 1,
 -1, -1, 1, 1,
 -1, 1, 1, -1
), byrow = TRUE, nrow = 3)
Cmat
```

```{r}
T.test <- function(X, Cmu){
 X <- as.matrix(X)
 n <- nrow(X)
 p <- ncol(X)
 df2 <- n-p
 xbar <- colMeans(X)
 S <- cov(X)
 alpha <- 0.05
 if(df2 < 1L) stop("Need nrow(X) > ncol(X).")
 if(length(Cmu) != p) mu <- rep(Cmu[1], p)
 
 T2 <- n * t((Cmat %*% xbar) - Cmu) %*% 
 solve(Cmat %*% S %*% t(Cmat)) %*%
 ((Cmat %*% xbar) - Cmu)
 
 Fstat <- T2 / ((p-1) * (n-1) / df2)
 C2 <- qf(1-alpha, df1 = p-1, df2 = n-p+1) * (n - 1) * (p - 1) / (n-p+1)
 
 pval <- 1 - pf(Fstat, df1 = p-1, df2 = df2)
 
 data.frame(T2 = as.numeric(T2), 
 Fstat = as.numeric(Fstat),
 C2 = as.numeric(C2),
 df1 = p-1, 
 df2 = df2, 
 p.value = as.numeric(pval), 
 row.names="") }
```

```{r}
Cmu <- c(0, 0, 0) 
T.test(datasoal_5, Cmu)
```
Interpretasi:
Karena T2 > F hitung dan p-value < alpha=0.05. Maka dari ini bisa disimpulkan bahwa tolak H0 dimana cmu = 0. Dengan tingkat signifikansi 5%, dapat di simpulkan bahwa dari data terdapat minimal 1 perbedaan nilai cmiu pada hasil test 2 model psikologi kemampuan numerik 


```{r}
C1 <- matrix(Cmat[1,], ncol = 1)
C2 <- matrix(Cmat[2,], ncol = 1)
C3 <- matrix(Cmat[3,], ncol = 1)

S.ci <- function(data, c, n, level = 0.95){
  q <- length(c)
  xbar <- colMeans(data)
  S <- cov(data)
  if(nrow(S) != q) stop("Need length(c) == nrow(S)") 
  if(ncol(S) != q) stop("Need length(c) == ncol(S).")
  if(level <=0 | level >= 1) stop("Need 0 < level < 1.")
 
  cval <- qf(level, q-1, n-q+1) * (q-1) * (n-1) / (n-q+1)
  
  # hitung z bar
  zhat <- t(c) %*% xbar
  
  # hitung varians z
  zvar <- (t(c) %*% S %*% c) / n
  
  # hitung standard error
  const <- sqrt(cval * zvar)
  
  # Membuat Smultaneous COnfidence Interval 
  c(lower = zhat - const, upper = zhat + const)
}

print(paste("CI untuk single contrast yang mewakili efek paritas"))

S.ci(datasoal_5, C1, n)

print(paste("CI untuk single contrast yang mewakili efek format"))

S.ci(datasoal_5, C2, n)

print(paste("CI untuk single contrast yang mewakili efek interaksi"))

S.ci(datasoal_5, C3, n)
```
Interpretasi:
1. Pada CI untuk single contrast yang mewakili efek paritas, tidak mengandung 0 yang artinya kesamaan paritas membutuhkan sedikit waktu untuk di jawab
2. Pada CI untuk single contrast yang mewakili efek format, tidak mengandung 0 yang artinya format arab membutuhkan sedikit waktu untuk di jawab
3. Pada CI untuk single contrast yang mewakili efek interaksi, mengandung 0 yang artinya tidak terdapat efek interaksi

Diketahui:
1. Model M: Ketiadaan interaksi
2. Model C: Kehadiran interaksi
Dapat dilihat pada hasil CI single contrast yang mewakili efek interaksi hasil nya mengansung 0. Artinya tidak terdapat efek interaksi. Oleh karena itu, ketiadaan interaksi itu mendukung model kognisi numerik M.

```{r}
#ubah data jadi matrix
matrix_data <- as.matrix(datasoal_5)
matrix_data
```
```{r}
# Difference score from the parity type contrast
diff_parity <- matrix_data %*% C1

# Difference score from the number type contrast
diff_format <- matrix_data %*% C2

# Difference score from the interaction type contrast
diff_interaction <- matrix_data %*% C3

diff_data <- data.frame(diff_parity, diff_format, diff_interaction)
diff_data
```
Cek Multivariate Normal
```{r}
mvn(data = diff_data[, c(1,2,3)], multivariatePlot = 'plot', showOutliers = TRUE)
```

```{r}
mshapiro.test(diff_data)
```
Cek outlier
```{r}
mvn(diff_data, multivariatePlot = "qq", showOutliers = TRUE)
```
Interpretasi:
Berdasarkan hasil dengan metode Henze-Zirkler dan shapiro wilk bahwa data tidak multivariate normal

