---
title: "Tugas pertemuan-10"
author: "Riofebri Prasetia (221911192)"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 12.8
Using the distances in Example 12.4, cluster the items using the average linkage hierarchical procedure. Draw the dendrogram. Compare the results with 
those in Examples 12.4 and 12.6.
```{r}
# dd <- dist(scale(USArrests), method = "euclidean")
dd <- matrix(c(0,9,3,6,11,
               9,0,7,5,10,
               3,7,0,9,2,
               6,5,9,0,8,
               11,10,2,8,0),5,5)
dd <- as.dist(dd, diag = TRUE)
#dd <- dist(dd, method = "euclidean")
dd
```
```{r}
#single linkage
hc <- hclust(dd, method = "single")
plot(hc)
```
```{r}
#complete linkage
hc2 <- hclust(dd, method = "complete")
plot(hc2)
```
## 12.10
Use Ward's method to cluster the 4 items whose measurements on a single variable X are given in the following table.
```{r}
# load data
x <- c(2,1,5,8)
x_measurements <- data.frame(x)
x_measurements
```
a)
```{r}
dd_measurement <- dist(x_measurements, method = "euclidean", diag = TRUE)
dd_measurement

```
```{r}
# use ward's method
hcMeasurement <- hclust(dd_measurement, method = "ward.D")
plot(hcMeasurement)
```
```{r}
#function join cluster
joint_cluster <- function(join){
    joinc <- as.matrix(join)
    meanc <- mean(joinc)
    sum <- 0
    for (i in 1:nrow(joinc)) {
      sum <- sum + (joinc[i]-meanc)^2 
    }
    return(sum)
}
```

```{r}
joint_cluster(rbind(2))
joint_cluster(rbind(1))
joint_cluster(rbind(5))
joint_cluster(rbind(8))
```
Bisa dilihat jika hanya terdapat 1 titik maka tidak terdapat ESS atau jarak

b)
```{r}
#data frame ESS
j_12 <- joint_cluster(c(2,1))
j_13 <- joint_cluster(c(2,5))
j_14 <- joint_cluster(c(2,8))
j_23 <- joint_cluster(c(1,5))
j_24 <- joint_cluster(c(1,8))
j_34 <- joint_cluster(c(5,8))

clusterEss <- matrix(c(j_12,j_13,j_14,0,0,0,
                       0,0,0,j_23,j_24,0,
                       0,0,0,0,0,j_34),6,3)

clusterEss <- as.data.frame(clusterEss)
clusterEss
```
C)
```{r}
dd_measurement <- dist(clusterEss, method = "euclidean", diag = TRUE)
dd_measurement
hcMeasurement <- hclust(dd_measurement, method = "ward.D")
plot(hcMeasurement)
```

## 12.12 
Repeat Example 12.12, starting with the initial groups (AC) and (BD) . Compare your solution with the solution in the example. Are they the same? Graph the items in terms of their ( x1 , x2) coordinates, and comment on the solutions.
```{r}
#data
Item_12 <- c("A", "B", "C", "D")
x1_12 <- c(5,-1,1,-3)
x2_12 <- c(3,1,-2,-2)
tabel_12 <- data.frame(Item_12, x1_12, x2_12)
matrix_12 <- as.matrix(tabel_12[,2:3])
matrix_12
```
```{r}
# distance k_mean k=2
# function join cluster k-means
joint_kMeans <- function(join){
    n <- nrow(join)
    p <- ncol(join)
    sum <- matrix(c(0), ncol = p, nrow = 1)
    for (j in 1:p){
      sum[1,j] <- sum(join[,j])  
    }
    
    sum2 <- sum/n
    return(sum2)
}

#function square distances
squareDistance <- function(table){
  n <- ncol(table)
  sum <- 0
  for (i in 1:n){
    sum <- sum + (table[1,i]-table[2,i])^2
  }
  return(matrix(sum))
}

#function square distances matrix
squareDistanceMatrix <- function(ca,cb){
  n <- nrow(ca)
  hasil <- matrix(c(0), ncol = n, nrow = 1)
  for (i in 1:n) {
    hasil[,i] <- squareDistance(rbind(ca[i,],cb))
  }
  return(hasil)
}
```

```{r}
#matrix coordinate of centroid AC dan BD
j_ac <- joint_kMeans(matrix_12[c(1,3),])
j_bd <- joint_kMeans(matrix_12[c(2,4),])
rbind(j_ac, j_bd)
j_ac

#matrix coordinate of centroid [A,(AC)], [A,(BD)]
print(paste("matrix coordinate of centroid [A,(AC)], [A,(BD)]"))
squareDistance(rbind(matrix_12[1,],
                     joint_kMeans(matrix_12[c(1,3),])))

squareDistance(rbind(matrix_12[1,],
                     joint_kMeans(matrix_12[c(2,4),])))

#matrix coordinate of centroid [B,(AC)], [B,(BD)]
print(paste("matrix coordinate of centroid [B,(AC)], [B,(BD)]"))
squareDistance(rbind(matrix_12[2,],
                     joint_kMeans(matrix_12[c(1,3),])))

squareDistance(rbind(matrix_12[2,],
                     joint_kMeans(matrix_12[c(2,4),])))

#matrix coordinate of centroid [C,(AC)], [C,(BD)]
print(paste("matrix coordinate of centroid [C,(AC)], [C,(BD)]"))
squareDistance(rbind(matrix_12[3,],
                     joint_kMeans(matrix_12[c(1,3),])))

squareDistance(rbind(matrix_12[3,],
                     joint_kMeans(matrix_12[c(2,4),])))

#matrix coordinate of centroid [D,(AC)], [D,(BD)]
print(paste("matrix coordinate of centroid [D,(AC)], [D,(BD)]"))
squareDistance(rbind(matrix_12[2,],
                     joint_kMeans(matrix_12[c(1,3),])))

squareDistance(rbind(matrix_12[2,],
                     joint_kMeans(matrix_12[c(2,4),])))


#square distance to group centroids A dan (BCD)
print(paste("square euclidian contoh A dan (BCD)"))
rbind(squareDistanceMatrix(matrix_12, matrix_12[1,]), squareDistanceMatrix(matrix_12, joint_kMeans(matrix_12[2:4,])))
```
dari hasil di atas tidak ada cluster yang dipindahkan sehingga coordinat centroid tidak diperbarui. Solusinya seperti contoh 12.12 yaitu dengan memulai dari inisial group (A,B) dan (C,D)

## 12.16
The national track records data for women are given in Table 1 .9. 
```{r}
#input data
url <- "https://raw.githubusercontent.com/rii92/tugas-APG/main/tugas%20pertemuan%2010/National%20Track%20Records%20for%20Women.csv"
ntrwomen <- read.csv(url)
mntrwomen <- as.matrix(ntrwomen[,2:7])
ntrwomen
```

(a) Using the data in Table 1.9, calculate the Euclidean distances between pairs 
of countries. 
```{r}
#nntrwoman <- nrow(mntrwomen)
#msntrwoman <- matrix(c(0), nrow = nntrwoman, ncol = nntrwoman)

#for (i in 1:nntrwoman) {
#  msntrwoman[i,] <- squareDistanceMatrix(mntrwomen, mntrwomen[i,])  
#}
#msntrwoman

ddntrwoman <- dist(mntrwomen, method = "euclidean", diag = TRUE)
as.matrix(ddntrwoman)
```

(b) Treating the distances in (a) as measures of (dis)similarity, cluster the countries using the single linkage and complete linkage hierarchical procedures. Construct dendrograms and compare the results.
```{r}
# dengan single linkage
hc2ntrwomen <- hclust(ddntrwoman, method = "single")
plot(hc2ntrwomen)

#dengan complete linkage
hc2ntrwomen2 <- hclust(ddntrwoman, method = "complete")
plot(hc2ntrwomen2)

```

(c) Input the data in Table 1.9 into a K-means clustering program. Cluster the countries into groups using several values of K. Compare the results with those in Part b.
```{r}
library(tidyverse)
library(cluster)
library(factoextra)
library(gridExtra)

kmeans2 <- kmeans(mntrwomen, centers = 2, nstart = 25)
kmeans3 <- kmeans(mntrwomen, centers = 3, nstart = 25)
kmeans4 <- kmeans(mntrwomen, centers = 4, nstart = 25)
kmeans5 <- kmeans(mntrwomen, centers = 5, nstart = 25)

str(kmeans2)
str(kmeans3) # to see the variables that are used
str(kmeans4)
str(kmeans5)


plot1 <- fviz_cluster(kmeans2, geom = "point", data = mntrwomen) + 
ggtitle("k = 2")
plot2 <- fviz_cluster(kmeans3, geom = "point", data = mntrwomen) + 
ggtitle("k = 3")
plot3 <- fviz_cluster(kmeans4, geom = "point", data = mntrwomen) + 
ggtitle("k = 4")
plot4 <- fviz_cluster(kmeans5, geom = "point", data = mntrwomen) + 
ggtitle("k = 5")
grid.arrange(plot1, plot2, plot3, plot4, nrow = 2)
```
perbandingan dengan metode hierarchical dengan single linkage dan complete linkage dengan metode k-means. pada metode hierarchical bisa dilihat tingkatan cluster secara detail di banding metode k-means. hal ini masih bisa ditangani karena jumlah observasi yang masih terhitung sedikit untuk metode hierarchical
