---
title: "APG_Pertemuan_6"
author: "Riofebri Prasetia (221911192)"
date: "3/23/2022"
output: html_document
---

```{r library, message=FALSE}
library('MVN')
library('mvoutlier')
library('rcompanion')
library('caret')
library('RVAideMemoire')
library('ellipse')
library('qqplotr')
library('evd')
library('outliers')
library('ggplot2')
library('SciViews')
library('MVTests')
```

6.7
Using the summary statistics for the electricity-demand data given in Example 6.4, compute T^2 and test the hypothesis H0: miu1 - miu2 = 0, assuming that sigma1(sum1) = sigma2(sum2) Set a = .05. Also, determine the linear combination of mean components most responsible for the rejection of H0

```{r}
# Inisiasi summary electricity-demand
alpha <- 0.05
p <- 2
n_1 <- 45
n_2 <- 55
x_mean_1 <- matrix(c(204.4,556.6), ncol = 1, nrow = 2)
x_mean_2 <- matrix(c(130.0, 355.0), ncol = 1, nrow = 2)
S_1 <- matrix(c(13825.3, 23823.4, 23823.4, 73107.4), 2,2)
S_2 <- matrix(c(8632.0, 19616.7, 19616.7, 55964.5), 2,2)
```

```{r}
# Function S_pooled
S_pooled <- function(n1, n2, S1, S2){
  S_pooled1 <- ((n1-1)/(n1+n2-2))*S1 + ((n2-1)/(n1+n2-2))*S2
  S_pooled1
}
```

```{r}
# S_pooled electricity-demand
S_pooled_ed <- S_pooled(n_1, n_2, S_1, S_2)
S_pooled_ed <- as.matrix(S_pooled_ed)
S_pooled_ed
```
```{r}
#hitung T_kuadrat
T_kuadrat <- t(x_mean_1 - x_mean_2)%*%(solve((1/n_1+1/n_2)*S_pooled_ed))%*%(x_mean_1 - x_mean_2)
T_kuadrat
```
```{r}
#hitung C_kuadrat
C_kuadrat <- ((n_1+n_2-2)*p/(n_1 + n_2 - p - 1))*qf(1-alpha, df1 = p, df2 = n_1+n_2-p-1)
C_kuadrat
```

```{r}
#perbandingan x mean yang paling berpengaruh 
perbandingan <- solve(S_pooled_ed)%*%(x_mean_1 - x_mean_2)
perbandingan
```
Interpretasi:
Berdasarkan hasil pada ringkasan example 6.4 maka kesimpulan adalah tolak H0 dimana H0: miu1-miu2 = 0. Artinya terdapat perbedaan pada kedua x mean. Kemudian pada hasil perbandingan x_mean yang paling berpengaruh untuk menolak H0 yaitu x mean 1 atau ukuran total konsumsi listrik pada puncaknya.

6.8 Observation on two response are collected for three treatments. The observation vectors t([x1, x2]) are

Treatment 1: t([6,7]), t([5,9]), t([8,6]), t([4,9]), t([7,9])

Treatment 2: t([3,3]), t([1,6]), t([2,3])

Treatment 3: t([2,3]), t([5,1]), t([3,1]), t([2,3])

a). Break up the observations into mean, treatment, and residual  components, as in (6-35). Construct the corresponding arrays for each variable. (See Example 6.8.)

b). Using the information in Part a, construct the one-way MANOVA table.

c). Set a = .01. Repeat the test using the chi-square approximation with Bartlett's correction. [See (6-39).] Compare the conclusions.

