---
title: "Tugas pertemuan-11"
author: "Riofebri Prasetia (221911192)"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(expm)
library(CCA)
```

## 10.10
In a study of poverty, crime, and deterrence, Parker and Smith [10] report certain summary crime statistics in various states for the years 1970 and 1973. A portion of their sample correlation matrix is

```{r}
ro11 <- matrix(c(1, .615, 
                .615, 1),2,2)
ro12 <- matrix(c(-.111, -.195, 
                -.266, -.085),2,2)
ro21 <- matrix(c(-.111, -.266, 
                -.195, -.085),2,2)
ro22 <- matrix(c(1, -.269, 
                -.269, 1),2,2)
```

The variables are 
$X_{1}^{(1)}$ == 1973 nonprimary homicides 
$X_{2}^{(1)}$ == 1973 primary homicides (homicides involving family or acquaintances) 
$X_{1}^{(2)}$ == 1970 severity of punishment (median months served) 
$X_{2}^{(2)}$ == 1970 certainty of punishment (number of admissions to prison 
divided by number of homicides)

(a) Find the sample canonical correlations.
```{r}
ro11s <- sqrtm(solve(ro11))
ro12s <- sqrtm(solve(ro12))
ro21s <- sqrtm(solve(ro21))
ro22s <- sqrtm(solve(ro22))

ro22i <- solve(ro22)
ro21i <- solve(ro21)
ro12i <- solve(ro12)
ro11i <- solve(ro11)

hasil <- ro11s%*%ro12%*%ro22i%*%ro21%*%ro11s
hasil
```

```{r}
eigen <- eigen(hasil)

eigenvector <- eigen$vectors
eigenvalues <- eigen$values

eigenvector
eigenvalues
```

```{r}
a <- ro11s%*%eigenvector
a
```

```{r}
b <- ro22i%*%ro21%*%a
b
```
```{r}
t(b)%*%ro22%*%b
sqrt(t(b)%*%ro22%*%b)
acc <- 1/(sqrt(t(b)%*%ro22%*%b))
b_final <- (acc[1,1])*b
b_final
```

```{r}
a
b_final
```
```{r}
ro1_bintang <- sqrt(eigenvalues[1])
ro2_bintang <- sqrt(eigenvalues[2])
ro1_bintang
ro2_bintang
```
hasil dari canonical correlation yaitu
$\rho_{1}^* = 0.327$ 
$\rho_{2}^* = 0.171$ 

(b) Determine the first canonical pair U1, V1 and interpret these quantities.
```{r}
a[,1]
b_final[,1]
```
Didapat modelnya ialah:
$U1 = a_{1}'Z^{(1)} = -1.00Z_{1}^{(1)} + .002Z_{2}^{(2)}$
$V1 = a_{1}'Z^{(1)} = -.611Z_{1}^{(1)} + .679Z_{2}^{(2)}$


## 10.11
Example 8.5 presents the correlation matrix obtained from n = 100 successive weekly rates of return for five stocks. Perform a canonical correlation analysis with $X^{1}=[X^{(1)}_{1}, X^{(1)}_{2}, X^{(1)}_{3}]$, the rates of return for the chemical companies, and $X^{1}=[X^{(1)}_{1}, X^{(1)}_{2}]$, the rates of return for the oil companies.

```{r}
rho11 <- matrix(c(1, .577, .509, 
                  .577, 1, .599,
                  .509, .599, 1),3,3)
rho12 <- matrix(c(.387,.389,.436,
                  .463,.322,.426),3,2)

rho21 <- matrix(c(.387,.462,
                  .389,.322,
                  .436,.426),2,3)

rho22 <- matrix(c(1,.532,.532,1),2,2)

rho11
rho12
rho21
rho22
```
```{r}
rho11s <- sqrtm(solve(rho11))
rho22s <- sqrtm(solve(rho22))

rho22i <- solve(rho22)
rho11i <- solve(rho11)

hasil_2 <- rho11s%*%rho12%*%rho22i%*%rho21%*%rho11s
hasil_2
```
```{r}
eigen2 <- eigen(hasil_2)

eigenvector2 <- eigen2$vectors
eigenvalues2 <- eigen2$values

eigenvalues2
eigenvector2
```
```{r}
a_2 <- rho11s%*%eigenvector2
a_2
```

```{r}
b_2 <- rho22i%*%rho21%*%a_2
b_2
```
```{r}
t(b_2)%*%rho22%*%b_2
sqrt(t(b_2)%*%rho22%*%b_2)
acc_2 <- 1/(sqrt(t(b_2)%*%rho22%*%b_2))
b_final2 <- (acc_2[1,1])*b_2
b_final2
```

```{r}
a_2
b_final2
```
```{r}
rho1_bintang <- sqrt(eigenvalues2[1])
rho2_bintang <- sqrt(eigenvalues2[2])
rho1_bintang
rho2_bintang
```

## 11.8
Use the diabetes data of Table 3.5. 
(a) Find the canonical correlations between (y1, y2) and (x1, X2, x3). 
(b) Find the standardized coefficients for the canonical variates. 
(c) Test the significance of each canonical correlation.
```{r}
url <- "https://raw.githubusercontent.com/rii92/tugas-APG/main/tugas%20pertemuan%2011/table%203.5.csv"
data3 <- read.csv(url)
data3 <- data3[,-1]
data3
```
a)
```{r}
cordata3 <- cor(data3)
cordata3
```
```{r}
y1y2 <- cordata3[c(1,2),c(1,2)]
x1x2x3 <- cordata3[c(3,4,5),c(3,4,5)]
rhoo11 <- y1y2
rhoo22 <- x1x2x3
rhoo12 <- cordata3[c(1,2), c(3,4,5)]
rhoo21 <- cordata3[c(3,4,5), c(1,2)]
rhoo11
rhoo22
rhoo12
rhoo21
```
```{r}
rhoo11s <- sqrtm(solve(rhoo11))
rhoo22s <- sqrtm(solve(rhoo22))

rhoo22i <- solve(rhoo22)
rhoo11i <- solve(rhoo11)

hasil3 <- rhoo11s%*%rhoo12%*%rhoo22i%*%rhoo21%*%rhoo11s
hasil3
```

```{r}

eigen3 <- eigen(hasil3)

eigenvector3 <- eigen3$vectors
eigenvalues3 <- eigen3$values

eigenvector3

eigenvalues3
```

```{r}
a_3 <- rhoo11s%*%eigenvector3
a_3
```
```{r}
b_3 <- rhoo22i%*%rhoo21%*%a_3
b_3
```
```{r}
acc3_1 <- 1/(sqrt(t(b_3[,1])%*%rhoo22%*%b_3[,1]))
acc3_2 <- 1/(sqrt(t(b_3[,2])%*%rhoo22%*%b_3[,2]))
b_final3_1 <- (acc3_1[1,1])*b_3[,1]
b_final3_2 <- (acc3_2[1,1])*b_3[,2]
b_final3_1
b_final3_2
```

```{r}
# a dan b
a_3
b_final3
```
```{r}
rhoo1_bintang <- sqrt(eigenvalues3[1])
rhoo2_bintang <- sqrt(eigenvalues3[2])
rhoo1_bintang
rhoo2_bintang
```
hasil dari canonical correlation seperti di atas

b)
```{r}
S_data3 <- cov(data3)
Sy_data3 <- S_data3[c(1,2),c(1,2)]
Sx_data3 <- S_data3[c(3,4,5),c(3,4,5)]
Sy_data3
Sx_data3
```
```{r}
asmatrixxdiag <- function(matrix){
  n <- nrow(matrix)
  p <- ncol(matrix)
  matrix_diag <- matrix(c(0), n, p)
  for (i in 1:n){
    matrix_diag[i,i] <- matrix[i,i]
  }
  return(matrix_diag)
}
```


```{r}
Dy <- asmatrixxdiag(Sy_data3)
Dx <- asmatrixxdiag(Sx_data3)
cy <- Dy%*%a_3
cx <- Dx%*%b_final3
cy
cx
```
Standardized coefficients untuk 2 canonical variates seperti di atas

c)
Uji Signifikansi Korelasi Kanonik
Ho: ðœ®12 = 0
Ha: ðœ®12 â‰  0
```{r}
n = 39
p = 2
q = 3

# Ho: rho12 = 0 vs Ha: rho12 != 0
stat.hit <- -(n-1-0.5*(p+q+1))*log((1-rhoo1_bintang^2) * (1-rhoo2_bintang^2))

stat.uji <- qchisq(0.95, p*q)

if(stat.hit > stat.uji){
  cat(stat.hit, ">", stat.uji, ", maka: ")
  cat("Tolak Ho")
}else{
  cat(stat.hit, "<", stat.uji, ", maka: ")
  cat("Gagal Tolak Ho")
}
```

```{r}
# Uji Ho: rho1 != 0 dan rho2 = 0
stat.hit <- -(n-1-0.5*(p+q+1))*log(1-rhoo2_bintang^2)

stat.uji <- qchisq(0.95, (p-1)*(q-1))

if(stat.hit > stat.uji){
  cat(stat.hit, ">", stat.uji, ", maka: ")
  cat("Tolak Ho")
}else{
  cat(stat.hit, "<", stat.uji, ", maka: ")
  cat("Gagal Tolak Ho")
}
```
Karena gagal tolak Ho, maka Rho1 atau korelasi kanonik yang pertama dapat dikatakan signifikan, sedangkan rho2 atau korelasi kanonik yang kedua tidak signifikan. Sehingga dalam analisis korelasi kanonik ini cukup hanya membentuk satu pasang set variabel saja.

## 11.9
Use the sons data of Table 3.8. 
```{r}
url <- "https://raw.githubusercontent.com/rii92/tugas-APG/main/tugas%20pertemuan%2011/table%203.8.csv"
data4 <- read.csv(url)
data4
```

(a) Find the canonical correlations between (y1,y2) and (x1,x2)Â· 
(b) Find the standardized coefficients for the canonical variates. 
(c) Test the significance of each canonical correlation.

a)
```{r}
cordata4 <- cor(data4)
cordata4
```
```{r}
rhooo11 <- cordata3[c(1,2),c(1,2)]
rhooo22 <- cordata3[c(3,4),c(3,4)]
rhooo12 <- cordata3[c(1,2), c(3,4)]
rhooo21 <- cordata3[c(3,4), c(1,2)]
rhooo11
rhooo22
rhooo12
rhooo21
```
```{r}
rhooo11s <- sqrtm(solve(rhooo11))
rhooo22s <- sqrtm(solve(rhooo22))

rhooo22i <- solve(rhooo22)
rhooo11i <- solve(rhooo11)

hasil4 <- rhooo11s%*%rhooo12%*%rhooo22i%*%rhooo21%*%rhooo11s
hasil4
```

```{r}
eigen4 <- eigen(hasil4)

eigenvector4 <- eigen4$vectors
eigenvalues4 <- eigen4$values

eigenvector4

eigenvalues4
```

```{r}
a_4 <- rhooo11s%*%eigenvector
a_4
```
```{r}
b_4 <- rhooo22i%*%rhooo21%*%a_4
b_4
```
```{r}
t(b_4)%*%rhooo22%*%b_4
sqrt(t(b_4)%*%rhooo22%*%b_4)
acc4 <- 1/(sqrt(t(b_4)%*%rhooo22%*%b_4))
b_final4 <- (acc4[1,1])*b_4
b_final4
```

```{r}
# a dan b
a_4
b_final4
```
```{r}
rhooo1_bintang <- sqrt(eigenvalues4[1])
rhooo2_bintang <- sqrt(eigenvalues4[2])
rhooo1_bintang
rhooo2_bintang
```
hasil dari canonical correlation seperti di atas

b)
```{r}
S_data4 <- cov(data4)
Sy_data4 <- S_data4[c(1,2),c(1,2)]
Sx_data4 <- S_data4[c(3,4),c(3,4)]
Sy_data4
Sx_data4
```
```{r}
asmatrixxdiag <- function(matrix){
  n <- nrow(matrix)
  p <- ncol(matrix)
  matrix_diag <- matrix(c(0), n, p)
  for (i in 1:n){
    matrix_diag[i,i] <- matrix[i,i]
  }
  return(matrix_diag)
}
```


```{r}
Dy4 <- asmatrixxdiag(Sy_data4)
Dx4 <- asmatrixxdiag(Sx_data4)
cy4 <- Dy4%*%a_4
cx4 <- Dx4%*%b_final4
cy4
cx4
```
Standardized coefficients untuk 2 canonical variates seperti di atas

c)
Uji Signifikansi Korelasi Kanonik
Ho: ðœ®12 = 0
Ha: ðœ®12 â‰  0
```{r}
n = 25
p = 2
q = 3

# Ho: rho12 = 0 vs Ha: rho12 != 0
stat.hit <- -(n-1-0.5*(p+q+1))*log((1-rhooo1_bintang^2) * (1-rhooo2_bintang^2))

stat.uji <- qchisq(0.95, p*q)

if(stat.hit > stat.uji){
  cat(stat.hit, ">", stat.uji, ", maka: ")
  cat("Tolak Ho")
}else{
  cat(stat.hit, "<", stat.uji, ", maka: ")
  cat("Gagal Tolak Ho")
}
```

```{r}
# Uji Ho: rho1 != 0 dan rho2 = 0
stat.hit <- -(n-1-0.5*(p+q+1))*log(1-rhooo2_bintang^2)

stat.uji <- qchisq(0.95, (p-1)*(q-1))

if(stat.hit > stat.uji){
  cat(stat.hit, ">", stat.uji, ", maka: ")
  cat("Tolak Ho")
}else{
  cat(stat.hit, "<", stat.uji, ", maka: ")
  cat("Gagal Tolak Ho")
}
```
Karena gagal tolak Ho, maka Rho1 atau korelasi kanonik yang pertama dapat dikatakan signifikan, sedangkan rho2 atau korelasi kanonik yang kedua tidak signifikan. Sehingga dalam analisis korelasi kanonik ini cukup hanya membentuk satu pasang set variabel saja.