---
title: "Tugas pertemuan-11"
author: "Riofebri Prasetia (221911192)"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(expm)
```

## 10.10
In a study of poverty, crime, and deterrence, Parker and Smith [10] report certain summary crime statistics in various states for the years 1970 and 1973. A portion of their sample correlation matrix is

```{r}
ro11 <- matrix(c(1, .615, 
                .615, 1),2,2)
ro12 <- matrix(c(-.111, -.195, 
                -.266, -.085),2,2)
ro21 <- matrix(c(-.111, -.266, 
                -.195, -.085),2,2)
ro22 <- matrix(c(1, -.269, 
                -.269, 1),2,2)
```

The variables are 
$X_{1}^{(1)}$ == 1973 nonprimary homicides 
$X_{2}^{(1)}$ == 1973 primary homicides (homicides involving family or acquaintances) 
$X_{1}^{(2)}$ == 1970 severity of punishment (median months served) 
$X_{2}^{(2)}$ == 1970 certainty of punishment (number of admissions to prison 
divided by number of homicides)

(a) Find the sample canonical correlations.
```{r}
ro11s <- sqrtm(solve(ro11))
ro12s <- sqrtm(solve(ro12))
ro21s <- sqrtm(solve(ro21))
ro22s <- sqrtm(solve(ro22))

ro22i <- solve(ro22)
ro21i <- solve(ro21)
ro12i <- solve(ro12)
ro11i <- solve(ro11)

hasil <- ro11s%*%ro12%*%ro22i%*%ro21%*%ro11s
hasil
```

```{r}
eigen <- eigen(hasil)

eigenvector <- eigen$vectors
eigenvalues <- eigen$values

eigenvector
```

```{r}
a1 <- ro11s%*%eigenvector[,1]
a1
```
```{r}
b1 <- ro22i%*%ro21%*%a1
b1
```
```{r}
t(b1)%*%ro22%*%b1
sqrt(t(b1)%*%ro22%*%b1)
acc <- 1/(sqrt(t(b1)%*%ro22%*%b1))
hasil_scale <- (acc[1,1])*b1
hasil_scale
```

```{r}
a1
hasil_scale
```
```{r}
ro1_bintang <- sqrt(eigenvalues[1])
ro2_bintang <- sqrt(eigenvalues[2])
ro1_bintang
ro2_bintang
```
hasil dari canonical correlation seperti di atas

(b) Determine the first canonical pair U1, V1 and interpret these quantities.
```{r}
a1
hasil_scale
```
$U1 = a_{1}$

didapat model nya ialah

```{r}
ro11 <- matrix(c(1, .4, .4, 1),2,2)
ro21 <- matrix(c(.5, .6, .3, .4),2,2)
ro12 <- matrix(c(.5,.3,.6,.4),2,2)
ro22 <- matrix(c(1,.2,.2,1),2,2)
```

## 10.11
Example 8.5 presents the correlation matrix obtained from n = 100 successive weekly rates of return for five stocks. Perform a canonical correlation analysis with $X^{1}=[X^{(1)}_{1}, X^{(1)}_{2}, X^{(1)}_{3}]$, the rates of return for the chemical companies, and $X^{1}=[X^{(1)}_{1}, X^{(1)}_{2}, X^{(1)}_{3}]$, the rates of return for the oil companies.

```{r}
rho11 <- matrix(c(1, .577, .509, 
                  .577, 1, .599,
                  .509, .599, 1),3,3)
rho12 <- matrix(c(.387,.389,.436,
                  .463,.322,.426),3,2)

rho21 <- matrix(c(.387,.462,
                  .389,.322,
                  .436,.426),2,3)

rho22 <- matrix(c(1,.532,.532,1),2,2)

rho11
rho12
rho21
rho22
```
```{r}
rho11s <- sqrtm(solve(rho11))
rho22s <- sqrtm(solve(rho22))

rho22i <- solve(rho22)
rho11i <- solve(rho11)

hasil_2 <- rho11s%*%rho12%*%rho22i%*%rho21%*%rho11s
hasil_2
```
```{r}
eigen2 <- eigen(hasil_2)

eigenvector2 <- eigen2$vectors
eigenvalues2 <- eigen2$values

eigenvalues2
eigenvector2
```
```{r}
a1_2 <- rho11s%*%eigenvector2[,1]
a1_2
```

```{r}
b1_2 <- rho22i%*%rho21%*%a1_2
b1_2
```
```{r}
t(b1_2)%*%rho22%*%b1_2
sqrt(t(b1_2)%*%rho22%*%b1_2)
acc_2 <- 1/(sqrt(t(b1_2)%*%rho22%*%b1_2))
hasil_scale_2 <- (acc_2[1,1])*b1_2
hasil_scale_2
```

```{r}
a1_2
hasil_scale_2
```
```{r}
rho1_bintang <- sqrt(eigenvalues2[1])
rho2_bintang <- sqrt(eigenvalues2[2])
rho1_bintang
rho2_bintang
```
